---
title: 'FDS Final Project: Report #3'
author: "Satu Iho"
output: html_document
---
```{r set_up, include=FALSE}
library(stringr)
library(httr)
library(purrr)
library(DT)
library(dplyr)
library(infer)
library(ggplot2)
library(kableExtra)
```

### Part 1. Example GET queries.

The highest grossing dramas in 2010 are shown in the table below.

```{r dramas_2010, echo=FALSE, message=FALSE, warning=FALSE}
grossing_dramas <-
httr::GET(url = stringr::str_glue("https://api.themoviedb.org/3/\\
                                  discover/movie?\\
                                  api_key=be7c9b4f37423cc11b204a90a2171479\\
                                  &primary_release_year=2010\\
                                  &sort_by=revenue.desc"))

grossing_dramas <- content(grossing_dramas, as="parsed") 

#Save first 5 results for display.
grossing_dramas <-
grossing_dramas$results %>% 
  map(purrr::pluck, "title") %>% 
  head(5)  

grossing_dramas <- tibble(grossing_dramas)

#Show in table
grossing_dramas %>% 
  rename(Title = grossing_dramas) %>% 
  kbl(caption = "5 highest grossing dramas in 2010") %>% 
  kable_paper(full_width = F)

```

The films that Will Ferrell and Liam Neeson have acted in together are shown in
the table below.

```{r will_liam, echo=FALSE, message=FALSE, warning=FALSE}
will_liam <- httr::GET(url = stringr::str_glue("https://api.themoviedb.org/3/\\
                              discover/movie?\\
                              api_key=be7c9b4f37423cc11b204a90a2171479\\
                              &with_people=23659,3896\\
                              &sort_by=vote_average.desc"))

will_liam <- content(will_liam, as="parsed") 

will_liam <- 
  will_liam$results %>% 
  map(purrr::pluck, "title")

will_liam <- tibble(will_liam)

#Show in table
will_liam %>% 
  rename(Title = will_liam) %>% 
  kbl(caption="Films with Will Ferrell and Liam Neeson") %>% 
  kable_paper(full_width = F)
```

```{r kids_tom, message=FALSE, warning=FALSE, include=FALSE}
tom_kids <- httr::GET(url = stringr::str_glue("https://api.themoviedb.org/3/\\
                              discover/movie?\\
                              api_key=be7c9b4f37423cc11b204a90a2171479\\
                              &certification_country=US&certification.lte=G\\
                              &with_cast=500"))

tom_kids <- content(tom_kids, as="parsed") 

tom_kids <- 
  tom_kids$results %>% 
  map(purrr::pluck, "title")

tom_kids <- tibble(tom_kids)

tom_kids_movies <-
tom_kids %>% 
  nrow()
```
There are **`r tom_kids_movies`** kids' movies with Tom Cruise in them.

### Part 2. Searcing for the Pixar company ID

```{r pixar_id, message=FALSE, warning=FALSE, include=FALSE}

pixar_id <- GET(str_glue("https://api.themoviedb.org/3\\
                                     /search/company?\\
                                     api_key=be7c9b4f37423cc11b204a90a2171479\\
                                     &query=%22pixar%22&page=1"))

pixar_id <- content(pixar_id, as="parsed") 

pixar_id <- 
pixar_id$results %>% 
  map_int(purrr::pluck, "id")    #Use map_int to return an integer, not a list.
```

The Pixar ID is: **`r pixar_id`**

### Part 3. Searching for Pixar films

The code in this section performs a query for all Pixar films and sorts
them in descending order in terms of revenue. The table then provides the film 
name and popularity as visible attributes for the queried list. 

```{r pixar_films, echo=FALSE}

#Set final page number to 6 as per total Pixar film count (113).
pages_pixar <- c(1:6)  

#Get data.
pixar_films <- str_glue("https://api.themoviedb.org/3/\\
                              discover/movie?\\
                              api_key=be7c9b4f37423cc11b204a90a2171479\\
                              &language=en-US&sort_by=revenue.desc\\
                              &page={pages_pixar}&with_companies={pixar_id}") %>% 
  map(GET) %>%
  map(content)  %>%
  map(purrr::pluck, "results") %>%
  purrr::flatten() %>%
  map_dfr(magrittr::extract, c("title","popularity"))

#Display the results in a nice format.
datatable(pixar_films)

```

### Part 4. Comparing Pixar and Walt Disney films.  

This section performs a comparison of Pixar and Disney films as per Disney's 
acquiring of Pixar in 2006. The analysis includes all films for each
company, with more than 50 votes, using *01.06.2006* as the cutoff date to 
distinguish between for Pixar (prior 01.06.2016) and Disney (post 01.06.2016).
The film *Cars* is therefore the first Pixar film classified as a Disney film,
as it was the first film to have been released under the [Disney-Pixar branding](https://en.wikipedia.org/wiki/Pixar) 

```{r disney_pixar_comparison, message=FALSE, warning=FALSE, include=FALSE}
#First step: get the Walt Disney Pictures ID and save it as a variable.
disney_id <- GET(str_glue("https://api.themoviedb.org/3/\\
                           search/company?\\
                           api_key=be7c9b4f37423cc11b204a90a2171479\\
                           &query=%22Walt%20Disney%20Pictures%22&page=1"))

disney_id <- content(disney_id, as="parsed") 


disney_id <- 
disney_id$results %>% 
   map_df(magrittr::extract, c("id", "name", "origin_country")) %>% 
   filter(name == "Walt Disney Pictures") %>% 
   pull(id)

#Second step: search for Disney films after 01.01.2006
pages_disney <- c(1:7)
disney_votes <- str_glue("https://api.themoviedb.org/3/\\
                              discover/movie?\\
                              api_key=be7c9b4f37423cc11b204a90a2171479\\
                              &language=en-US&sort_by=vote_cound.desc\\
                              &page={pages_disney}\\
                              &primary_release_date.gte=2006-06-01\\
                              &vote_count.gte=50\\
                              &with_companies={disney_id}") %>% 
  map(GET) %>%
  map(content)  %>%
  map(purrr::pluck, "results") %>%
  purrr::flatten() %>%
  map_dfr(magrittr::extract,
          c("title",
            "vote_average",
            "vote_count",
            "release_date"))

#Second sub-step: get all Pixar films before 01.01.2006.
pixar_votes <- str_glue("https://api.themoviedb.org/3/\\
                              discover/movie?\\
                              api_key=be7c9b4f37423cc11b204a90a2171479\\
                              &language=en-US&sort_by=vote_cound.desc\\
                              &page={pages_pixar}\\
                              &primary_release_date.lte=2006-06-01\\
                              &vote_count.gte=50\\
                              &with_companies={pixar_id}") %>% 
  map(GET) %>%
  map(content) %>%
  map(purrr::pluck, "results") %>%
  purrr::flatten() %>%
  map_dfr(magrittr::extract,
          c("title",
            "vote_average",
            "vote_count",
            "release_date"))

#Add a company identifier and combine the Disney and Pixar votes tables. 
pixar_votes <-
pixar_votes %>% 
  mutate(company = "Pixar")

disney_votes <-
disney_votes %>% 
  mutate(company = "Disney")

all_votes <-
pixar_votes %>% 
  bind_rows(disney_votes)

#The API seems to bring in a lot of duplicates, so I'll remove those.
all_votes <-
all_votes %>% 
  distinct()

#Have a first glance at the data (not shown in output, only used during 
#initial analysis and as a sense check)
all_votes %>% 
  group_by(company) %>% 
  summarise(count = n(),
            median = median(vote_average),
            mean = round(mean(vote_average), digits=2),
            standard_deviation = round(sd(vote_average),digits=2),
            min = min(vote_average),
            max = max(vote_average)
            )
```

First, let's look at a boxplot for the two companies' vote averages.

```{r disney_pixar_comparison_plot, echo=FALSE, message=FALSE, warning=FALSE}
#Draw boxplot to view the data. 
all_votes %>% 
  group_by(company) %>% 
  ggplot(aes(x=company, y=vote_average)) +
  geom_boxplot(alpha = 0.1, fill="blue")+
  coord_flip()+
  theme_classic()+
  scale_y_continuous(breaks = c(5, 5.5, 6, 6.5, 7, 7.5, 8))+
  labs(title="Comparing Disney and Pixar's vote averages",
       x="Company", y="Vote average",
       caption= "Data from The Movie Database")

```

Based on the plot it seems like films produced under Disney's ownership have 
been less popular with the public than the original Pixar films, with a lower 
median and the votes being skewed to the lower end of the scale for Disney. 

Next, let's do a formal t-test to test for a statistically significant 
difference in the two vote average means.

```{r t_test_disney_pixar, echo=FALSE, message=FALSE, warning=FALSE}
#Run the t-test
t_test_votes <-
all_votes %>% 
  group_by(company) %>% 
  t_test(formula = vote_average ~ company,
                   order = c("Disney", "Pixar"),
                   conf_level = 0.95)

t_test_votes %>% 
  kbl(caption = "t-test Disney - Pixar vote averages") %>% 
  kable_paper(full_width = F)
```

The t-test suggests that there is no statistically significant difference in the
vote means for the two companies. 


