---
title: "FDS Final Project: Report #1"
output: html_document
---

### Part 1 Scraping the raw data and preparing it for analysis.  

*No output is shown here on purpose.*

```{r setup, message=FALSE, warning=FALSE, include=FALSE}

#Basic setup
library(httr) 
library(rvest)
library(stringr)
library(dplyr)
library(tidyr)
library(lubridate)
library(digest)
library(styler)
library(ggmap)
library(remedy)
library(ggplot2)
library(infer)

#This part reads the raw html data into r.
flats_html <- xml2::read_html("https://epfl-exts.github.io/rental-scrape/")

#Scrape each data element
location <-
flats_html %>% 
  html_nodes(css = ".address") %>% 
  html_text()

rooms <-
flats_html %>% 
  html_nodes(css = ".rooms") %>% 
  html_text(trim = TRUE) %>% 
  str_replace(" Rooms", "")

living_space <-
flats_html %>% 
  html_nodes(css = ".living-space") %>% 
  html_text() %>% 
  str_replace("Living space: ", "")

floor <- 
flats_html %>% 
  html_nodes(css = ".floor") %>% 
  html_text(trim=TRUE) %>% 
  str_replace("Floor: ", "") 

flats_html %>% 
  html_node(css =".price") %>% 
  html_text()

object_type <-
flats_html %>% 
  html_nodes(css = ".object-type") %>% 
  html_text(trim = TRUE)

price <-
flats_html %>% 
  html_nodes(css = ".price") %>% 
  html_text(trim = TRUE)

usable_surface <-
  flats_html %>% 
  html_nodes(css = ".usable-surface") %>% 
  html_text(trim = TRUE) %>% 
  str_replace("Usable surface: ", "")

availability <- 
  flats_html %>% 
  html_nodes(css = ".availability") %>% 
  html_text(trim = TRUE) %>% 
  str_replace("\nSee More", "") %>% 
  str_replace("Availability: ", "")

#Combine elements into a tibble for further cleansing.
scraped_rental_raw <- 
  tibble(location, 
         price,
         object_type,
         rooms, 
         living_space, 
         floor,
         availability,
         usable_surface
  )
```


```{r clean_up, message=FALSE, warning=FALSE, include=FALSE}

#Re-organise the price data into price and currency columns
scraped_rental <-
  scraped_rental_raw %>% 
  mutate(price = if_else(price == "Price on request NA", "", price)) %>%  
  #Clear out the cells with no price.
  mutate(price = str_split(price, " ")) %>%   
  #Split the price into two strings: the price value and currency
  unnest_wider(price) %>%                     
  #Move the strings into their own, separate columns
  rename(price = ...1, currency = ...2) %>%   
  #Rename the newly created columns to price and currency respectively
  mutate(price = as.numeric(price))           
 #Convert the price column from character into numeric 
  
#Clean up and convert the living_area column to numeric
scraped_rental <-
  scraped_rental %>% 
  mutate(living_space = if_else(living_space == "NA", 
                                "", 
                                str_replace(living_space, "m2", ""))) %>% 
  mutate(living_space = as.numeric(living_space)) 
  
#Finally, convert the availability column data from character to date format...
scraped_rental<-
  scraped_rental %>% 
  mutate(availability = dmy(availability)) 

#...and convert the rooms column data into numeric
scraped_rental <- 
  scraped_rental %>% 
  mutate(rooms = as.numeric(rooms))
```

### Part 2. Plotted relationship between price and living space.  

This part plots visually the relationship between price and living space.   

```{r plot_price_living, echo=FALSE, warning=FALSE}

scraped_rental %>%
  ggplot(mapping=aes(x=living_space, 
                     y=price))+
  geom_point(alpha=0.4)+
  labs(title="Plotted relationship between price and living space",
       caption="Data from RentalAgency website",
       x="Living space in m2", 
       y="Price per month")

```

The price - living space relationship seems rather consistent up to 200m2 living
space, but weakens somewhat as living space exceeds 200m2. 

### Part 3. Displaying the listings by postcode

```{r chart_post_code, echo=FALSE, warning=FALSE}

#First, create a tibble for postcode from location column in scraped_rental.
post_code_data <- scraped_rental %>%
    select(location) %>%            
    mutate(post_code = as.numeric(str_extract_all(location, "[:digit:]{4}"))) %>% 
  #Look for and retain only 4 digit strings within the data. 
    select(post_code)                             

#Then, create a tibble with listings per postcode.
listings_per_post_code <- post_code_data %>% 
  count(post_code)

#Finally, draw the barchart 
listings_per_post_code %>%
  ggplot(mapping=aes(x=post_code, 
                     y=n)) +
  geom_col()+
  theme_classic()+
  labs(title="Listings per postcode",
       caption="Data from RentalAgency website",
       x="Postcode", 
       y="Number of listings")
```

The agency seems to be more active in the postcodes in the early 1200's.

### Part 4. Price of flats over living space and floor.

```{r facet_plot, echo=FALSE, warning=FALSE}
#First, create a tibble that includes also postcode from location column 
price_floor_postcode <- scraped_rental %>%
    select(location, floor, price, living_space) %>%            
    mutate(post_code = as.numeric(str_extract_all(location,"[:digit:]{4}"))) %>% 
    select(-location) %>% 
    mutate(floor = as.numeric(if_else(floor == "NA", "", floor))) %>% 
    filter(floor < 7) %>%             #Include only floors 1-6.
    mutate(post_code_categories =     
             #Add a column with postcode categories for visualisation below.
             case_when(                   
     post_code <= 1210 ~ "1200-1210",
     post_code <= 1220 ~ "1211-1220",
     post_code <= 1250 ~ "1221-1250",
     post_code <= 1260 ~ "1251-1260",
     TRUE ~ "1261-")
     )

#Set colours for the plot
colours <- c("#1F77B4", "#FF7F0E", "#2CA02C",
            "#D62728", "#E377C2")

price_floor_postcode %>% 
  ggplot(aes(x=living_space, y=price)) +
  facet_wrap(vars(floor))+
  geom_point(aes(color=post_code_categories))+ 
  theme(legend.position = "bottom")+  
  scale_color_manual(values = colours) +
  labs(title="Price of flats over living space",
       caption="Displaying only floors 1-6.",
       x="Surface in m2", y="Price in CHF")
```

As per the colours of the plot, it seems that some postcodes are more prominent
in having buildings with three or more floors. These are presumably the 
postcodes for city centres.  

### Part 5. Comparing listings with address and address on request.   

```{r address_or_0, warning=FALSE, include=FALSE}
#Prepare the data for this analysis
address_on_request <-
scraped_rental %>% 
  filter(object_type != "Hobby room" & object_type != "Cellar compartment") %>% 
  #Filter out outliers in terms of price 
  mutate(on_request = str_detect(location, "request|demande")) %>%  
  #Add a column indicating whether address is on request or not (TRUE/FALSE)
  select(on_request, location, price, living_space, floor) %>%    
  mutate(floor = as.numeric(if_else(floor == "NA", "", floor))) 
```

In this section, we compare listings whose address is available on request only
(Address on request = true) and that have an address provided already in the 
listing (Address on request = false).   

First, we view the two types of listings in terms of their price, observing that
listings with address available on request being generally more expensive than 
those with an address provided.   

```{r address_or_1, echo=FALSE, warning=FALSE}
address_on_request %>% 
  group_by(on_request) %>% 
  ggplot(aes(x=on_request, 
             y=price)) +
  geom_boxplot(alpha = 0.1, fill="red")+
  coord_flip()+
  theme_classic()+
  labs(title="Examining the prices of listings with and without address provided",
       x="Address on request", y="Price")
```

Second, we compare the two types of listings in terms of their living space, 
observing that listings with address available on request being generally larger
than those with an address provided.   
```{r address_or_2, echo=FALSE, warning=FALSE}
address_on_request %>% 
  group_by(on_request) %>% 
  filter(living_space > 10) %>%     #Remove potential living space outliers 
  ggplot(aes(x=on_request, 
             y=living_space)) +
  geom_boxplot(alpha = 0.1, fill="blue")+
  coord_flip()+
  theme_classic()+
  scale_y_continuous(breaks = 
                       c(50, 100, 150, 200, 250, 300, 350, 400, 450, 500))+
  labs(title="Examining the living spaces of listings with and without address provided",
       x="Address on request", y="Living space")
```

Finally, comparing in terms of floors, it seems that listings whose addresses are available on request only, tend to be located on lower floors. This makes sense as these listings are likely to be larger (as per previous graph on living space) and hence they might be for instance villas or houses rather than apartments.   
```{r address_or_3, echo=FALSE, warning=FALSE}

#For better visibility of the graph, the one floor outlier (floor: 50, which can be 
#presumed to be a data key-in error) is excluded.
address_on_request %>% 
  group_by(on_request) %>% 
  filter(floor < 20) %>%     
  ggplot(aes(x=on_request, 
             y=floor)) +
  geom_boxplot(alpha = 0.1, fill="green")+
  coord_flip()+
  theme_classic()+
  scale_y_continuous(breaks = c(1:20))+ #Enhance visibility of graph with a more detailed scale
  labs(title="Examining the floors of listings with and without address provided",
       x="Address on request", y="Floor")
```

### Part 6. Statistical comparison of listings with address and address on request, in terms of price per square meter.   

In this section, we display a statistical analysis of listings whose address is
available on request only (Address on request = true) and that have an address 
provided already in the listing (Address on request = false) **in terms of their
price per square meter.**  
```{r address_or_stats, echo=FALSE}
address_on_request %>% 
  filter(price > 100) %>%           #Remove price outliers
  filter(living_space > 10) %>%     #Remove living space outliers
  mutate(price_per_sq = price/living_space) %>% 
  group_by(on_request) %>% 
  summarise(count = n(),
            median = round(median(price_per_sq), digits=2),
            mean = round(mean(price_per_sq), digits=2),
            standard_deviation = round(sd(price_per_sq),digits=2),
            min = round(min(price_per_sq), digits=2),
            max = round(max(price_per_sq), digits=2)
  )
  
```
Initial impression: the means between the two groups don't actually seem very 
different from each other, as well as the medians and standard deviations are 
rather similar. Intuitively, this would suggest that there is no statistically 
significant difference between the two types of listings.    

Let's however review this closer with a formal t-test:  
```{r t_test_price_per_sq, echo=FALSE}
address_on_request %>% 
  filter(price > 100) %>%         
  filter(living_space > 10) %>%   
  mutate(price_per_sq = price/living_space) %>% 
  group_by(on_request) %>% 
  t_test(formula = price_per_sq ~ on_request,
                   order = c("TRUE", "FALSE"),
                   conf_level = 0.95)
```

The t-test's p-value suggests that there no statistically significant difference
between the mean price per square meter for listings with and without addresses.
This makes some sense as the price per square meter could be expected to be 
similar regardless of the size of a home.  

### Part 7. Statistical comparison of listings with address and address on request, in terms of price.   
In this section, we display a statistical analysis of listings whose address is 
available on request only (Address on request = true) and that have an address 
provided already in the listing (Address on request = false) **in terms of their
price.**    

```{r t_test_price, echo=FALSE}
address_on_request %>% 
  filter(price > 10) %>%          #Filter out rows with nonsensical prices
  group_by(on_request) %>% 
  t_test(formula = price ~ on_request,
                   order = c("TRUE", "FALSE"),
                   conf_level = 0.95)
```
This t-test p-value suggests that there is a statistically significant 
difference between the mean prices of the two types of listings (with and 
without address) when compared in terms of the price only. This makes sense as 
we observed before that the listings with no address listed seem to be more 
expensive.  

### Part 8. Plotting addresses as per their longitude and latitude.  

The below map shows a sample of listing locations from the RentalAgency website.

```{r plot_map_data, include=FALSE}
#Get longitude and latitude data using the Google geolocation API
register_google(key = "AIzaSyD4KWAm6FqMs-5k8ARR-mgyIbYwzscYaRk")
getOption("ggmap")

#Store the lon-lat data of a sample of 35 addresses into a tibble.
long_lat_addresses <-
address_on_request %>% 
 filter(on_request == "FALSE") %>%   #Exclude the rows where address is available on request only.
 sample_n(35) %>%                    #Take a sample of 35 addresses.
 mutate_geocode(location)            #Add columns with the lon and lat details of the address.
```


```{r plot_map, echo=FALSE, message=FALSE, warning=FALSE}
#I tried the ggmap at first. 
#vaud_map <- 
#get_stamenmap(bbox = c(left=5.9196, bottom=46.1073, right=6.3297, top=46.3497))

#ggmap(vaud_map) +
#geom_point(data=long_lat_addresses, 
            #aes(x=lon, y=lat), alpha=0.9, color="purple")+
            #labs(title="Where are RentalAgency listings in Vaud and Geneva cantons?",
            #subtitle="Viewing a sample of 30 listings")

#The leaflet map below seems however a lot prettier than the generated with the
#code above, so I'll opt to use the leaflet one.
library(leaflet)
long_lat_addresses %>%           
 leaflet() %>% 
   addTiles() %>% 
   addMarkers(lng=~lon,lat=~lat)
```

