---
title: "FDS Final Project: Report #4"
author: "Satu Iho"
output: html_document
always_allow_html: true
---

```{r setup, include=FALSE}
#Basic set-up to load libraries and data. 
library(dplyr)
library(stringr)
library(ggplot2)
library(tidylog)
library(tidyr)
library(janitor)
library(digest)
library(purrr)
library(lubridate)
library(styler)
library(readxl)
library(DT)
library(kableExtra)
library(infer)

#Read in the data
athletes  <- read_excel("olympics.xlsx", sheet = "athletes")
countries <- read_excel("olympics.xlsx", sheet = "country")
games     <- read_excel("olympics.xlsx", sheet = "games")
medals    <- read_excel("olympics.xlsx", sheet = "medals")

#Clean up column names
athletes  <- clean_names(athletes)
countries <- clean_names(countries)
games     <- clean_names(games)
medals    <- clean_names(medals)
```

### Part 1. Have some athletes competed for different countries over time?

```{r multiple_country_athletes, include=FALSE}

#Identify athletes who have represented more than one country.
multiple_country_athletes <- 
countries %>% 
  distinct(athlete_id, noc) %>% 
  count(athlete_id) %>% 
  filter(n > 1)

#Save overall number to variable to use in output.
n_multiple_country_athletes <-
multiple_country_athletes %>% 
  nrow()
```

Overall, there have been **`r n_multiple_country_athletes`** athletes that have
competed representing different countries. While at first glance the number 
seems very large, it can be partially explained by geopolitical events e.g. the 
German reunification, the collapse of the Soviet Union etc. where it was rather
the (NOC affiliated) country that changed, rather than the athlete for example 
relocating and taking a different nationality. The data set goes all the way 
back to 1896, leaving plenty of time for such historical geopolitical events to
occur, changing athletes' affiliated countries in the process.

### Part 2. Who are the ten athletes that took part in most games?

The below table displays the ten athletes that have participated in the most 
games over the years and their sports.

```{r top_ten_games, echo=FALSE, message=FALSE, warning=FALSE}

#Create a filter for athletes that appear the most times. Somewhat 
#counterintuitively the countries table seems like the best place to start as it
#has the athlete_id and games combined already.
top10_athletes_filter <-
countries %>% 
  group_by(athlete_id) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) %>% 
  head(10) %>% 
  pull(athlete_id)

#Bring in the name of the athletes.
top10_athletes_with_names <-
countries %>% 
  filter(athlete_id %in% top10_athletes_filter) %>% 
  left_join(athletes, by=c("athlete_id" = "id")) %>% 
  select(athlete_id, name, age, noc, games)

#Do a sense check by also looking at the sport of the top 10 athletes.
top10_athletes_with_names_and_sports <-
top10_athletes_with_names %>% 
  left_join(medals, by="athlete_id") %>% 
  select(name, age, noc, sport) %>% 
  distinct(name, sport) %>% 
  rename("Name" = name,
         "Sport" = sport)

#Display results.
top10_athletes_with_names_and_sports %>% 
  kbl() %>% 
  kable_classic(full_width = F)

```

### Part 3. What athlete(s) kept a Gold medal for the longest time?

The below table displays the **top 15 Gold medal streaks by individuals 
athletes**, i.e. individuals that won a Gold medal in the same event the most 
times in a row in consecutive games. 

```{r gold_streak_athletes, echo=FALSE, message=FALSE, warning=FALSE}

#Add an id for each games and previous game
games <- 
  games %>% 
  arrange(season, year) %>% 
  mutate(games_id = row_number())

#Retain only gold medal winning athletes.
medals_gold <-
  medals %>% 
  filter(medal == "Gold")  
  
#Join the two tables and removed unneeded columns.
games_medals <- 
  games %>% 
  full_join(medals_gold, by="games") %>%  
  select(-year, -season, -city, -team, -sport, -games)

#Group by athlete and event, and identify their previous games.
games_medals <-
  games_medals %>% 
  arrange(athlete_id) %>% 
  group_by(athlete_id, event) %>% 
  mutate(previous_games = lag(games_id, default = -99))

#Identify starts of streaks
games_medals <-
  games_medals %>% 
   mutate(streak_begin = games_id - previous_games != 1)  

#Add a streak_id to facilitate counting
games_medals <-
  games_medals %>% 
  mutate(streak_id = cumsum(streak_begin))

#Add a counter
games_medals_count <-
games_medals %>% 
  count(streak_id, sort=TRUE)

#Make the final data nicer by adding athlete names.
consecutive_gold_top_athletes <-
  games_medals_count %>% 
  head(15) %>% 
  left_join(athletes, by=c("athlete_id" = "id")) %>% 
  select(name, event, n)

#Prepare data to display
consecutive_gold_top_athletes <-
consecutive_gold_top_athletes %>% 
  rename("Name" = name, 
         "Event" = event,
         "# of consecutive golds" = n)

#Display data.
consecutive_gold_top_athletes %>% 
  kbl(caption = "Athletes with longest gold streak per event") %>% 
  kable_classic(full_width = F)
```

### Part 4. What country(ies) kept a Gold medal for the longest time?

The below table displays the **top 15 Gold medal streaks by country**, i.e. 
countries that won a Gold medal in the same event the most times in a row in 
consecutive games. 

```{r gold_streak_countries, echo=FALSE, message=FALSE, warning=FALSE}

#Reuse the two first tibbles created in the previous part. 
countries_games_medals <- 
  games %>% 
  full_join(medals_gold, by="games") %>% 
  distinct(games_id, team, event)

#Group by country (team) and event, and identify their previous games.
countries_games_medals <- 
  countries_games_medals %>% 
  arrange(team) %>% 
  group_by(team, event) %>% 
  mutate(previous_games = lag(games_id, default = -99))

#Identify starts of streaks
countries_games_medals <- 
  countries_games_medals %>% 
   mutate(streak_begin = games_id - previous_games != 1)  

#Add a streak_id to facilitate counting
countries_games_medals <- 
  countries_games_medals %>% 
  mutate(streak_id = cumsum(streak_begin))

#Add a counter
countries_games_medals_count <- 
  countries_games_medals %>% 
  count(streak_id, sort=TRUE)

#Prepare data to display
countries_games_medals_count <- 
  countries_games_medals_count %>% 
  head(15) %>% 
  select(team, event, n) %>% 
  rename("Country" = team, 
         "Event" = event,
         "# of consecutive golds" = n)

#Display data
countries_games_medals_count %>% 
  kbl(caption = "Countries with longest gold streak per event") %>% 
  kable_classic(full_width = F)
```

### Part 5. Who are the ten athletes that competed in the most events?

The below table displays athletes that took part in the largest number of events
(over potentially multiple games).

```{r athletes_most_events, echo=FALSE, message=FALSE, warning=FALSE}
athletes_most_events <-
medals %>% 
  distinct(athlete_id, event) %>% 
  count(athlete_id, sort=TRUE) %>% 
  head(10)

#Prepare data for display
athletes_most_events_display <-
  athletes_most_events %>% 
  left_join(athletes, by=c("athlete_id" = "id")) %>% 
  select(name, n) %>% 
  rename("Name" = name, 
         "# of events" = n)

#Display data
athletes_most_events_display %>% 
 kbl(caption = "Athletes with most events") %>% 
  kable_classic(full_width = F)
```

### Part 6. Medals per country.

The below table displays the top 15 countries in terms of overall medals won 
over the years.

```{r medals_per_country, echo=FALSE, message=FALSE, warning=FALSE}
#Start off with the medals table, include only entries with a medal.
medals_per_country <-
medals %>% 
  filter(medal != "") %>% 
  select(games, team, medal)

#Find top 15 countries in terms of appearances (i.e. medals)
medals_per_country_filter <-
medals_per_country %>% 
  arrange(team) %>% 
  count(team, sort=TRUE) %>% 
  head(15) %>% 
  pull(team)

#Apply filter of top 15 countries.
medals_per_country <-
medals_per_country %>% 
  filter(team %in% medals_per_country_filter)

#Count medals by country and games.
medals_per_country <-
medals_per_country %>% 
  mutate(medal = "medal") %>%  #Convert all medals to uniform name to count them
  group_by(team, games) %>% 
  count(medal)

medals_per_country <-
medals_per_country %>% 
  select(-medal) %>% 
  arrange(games) %>% 
  pivot_wider(names_from = "games",
              values_from = "n")

#Display data
medals_per_country %>% 
 kbl(caption = "Countries with most medals") %>% 
  kable_paper(full_width = F) %>% 
  scroll_box(width = "100%")

```

### Part 7. Is there are relationship between country and the probability of 
### winning a medal?

```{r country_medal_probability, echo=FALSE, message=FALSE, warning=FALSE}
#Calculate overall number of athletes that have competed in one or more events 
#per country.
athletes_per_country <-
medals %>% 
  filter(team %in% medals_per_country_filter) %>% 
  distinct(athlete_id, team, event) %>% 
  group_by(team) %>% 
  summarise(total_athletes = n())

#Calculate total medals per country
medals_per_country <-
medals_per_country %>% 
  mutate(total_medals = rowSums(dplyr::across(2:51), na.rm = T))

#Combine the two into a tibble and calculate percentage
medals_athletes_perc <-
  medals_per_country %>% 
  select(team, total_medals) %>% 
  left_join(athletes_per_country, by="team") %>% 
  mutate(ma_perc = total_medals/total_athletes)

#Draw the bar chart.
medals_athletes_perc %>% 
  ggplot(mapping = aes(   x = team,
                          y = ma_perc))+
  geom_col() +
  geom_text(aes(label = round(ma_perc, digits=2)), 
            position = position_stack(vjust = 0.8),
            color = "white")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs(title="Medal/athlete ratio by country",
       x="", 
       y="")
```

To me, the bar chart suggests that there is a relationship between country and 
winning a medal. The two countries that stand out are the Soviet Union and East
Germany, where it seems that athletes from the two countries were had a high
likelihood of coming away with a medal if they were sent to the Olympics by
their respective countries.

```{r country_medal_probability_chi, echo=FALSE, message=FALSE, warning=FALSE}
#Chi-square test
medals_chi <-
  medals %>% 
  filter(team %in% medals_per_country_filter) %>% 
  mutate(medal_category = case_when(
    medal != "" ~ "Medal",
    TRUE ~ "No medal"))

medals_chi <-
  medals_chi %>% 
  mutate(medal_category = factor(medal_category, levels = c("Medal", 
                                                            "No medal")))

medals_chi_result <-
chisq_test(medals_chi, 
           formula = medal_category ~ team)

medals_chi_result %>% 
  kbl(caption = "Relationship between country and medal/no medal") %>% 
  kable_paper(full_width = F)
```

The results of the chi-squared test suggest that we can reject the null 
hypothesis of: "There is no relationship between country and whether they win a
medal or not". and that there is indeed a relationship between country and
winning a medal. 

### Part 8. Average heights and weights per sport.

The below scatter plot displays average athlete heights and weights per sport.
It also uses average BMI to determine an individual dot's size. The minimum and
maximum (height, weight and BMI) values have been labeled. **(There are only 
five labels as rhythmic gymnastics accounts for both the min BMI and min weight.)**

```{r athletes_wh, echo=FALSE, fig.height=8, fig.width=13, message=FALSE, warning=FALSE}
#Get height and weight data
athletes_hw <-
  athletes %>% 
  select(id, height, weight) %>% 
  filter(height != "" & weight !="")

#Get sport data
athletes_sport <-
  medals %>% 
  select(athlete_id, sport) %>% 
  distinct(athlete_id, sport)

#Combine the two for a full data set to work with
athletes_hw_sport <- 
athletes_hw %>% 
  inner_join(athletes_sport, by=c("id"="athlete_id"))

#Summarise the values by sport.
athletes_hw_sport <-
athletes_hw_sport %>% 
  group_by(sport) %>% 
  summarise(avg_weight = mean(weight),
            avg_height = mean(height),
            avg_BMI = mean(weight/(height/100)^2))

#Draw the scatterplot
athletes_hw_sport %>% 
  ggplot(mapping=aes(x    =avg_weight, 
                     y    =avg_height,
                     size =avg_BMI,
                     color=sport))+
  geom_point()+
  geom_text(data = athletes_hw_sport %>% 
              filter(avg_height %in% range(avg_height) |
                     avg_weight %in% range(avg_weight)|
                     avg_BMI %in% range(avg_BMI)), 
            aes(label=sport), 
            nudge_x = 0.25, nudge_y = 1.1)+
  labs(title="Athlete athlete weights and heights, with BMI",
       subtitle="BMI is represented by the size of the scattered dots and text 
       label sizes",
       x="Weight", y="Height")
 
#Check the labels are correct.
labels <-
  athletes_hw_sport %>% 
  filter(avg_weight == max(avg_weight) |
           avg_weight == min(avg_weight) |
           avg_height == max(avg_height) |
           avg_height == min(avg_height) |
           avg_BMI == max(avg_BMI) |
           avg_BMI == min(avg_BMI))


```

### Part 9. Number of medals given by year.

The below table displays the evolution of the number of medals for both the 
summer and winter games. While proportionally fewer medals are given out in the 
winter games (presumably simply because there are fewer events, especially team
events), both totals have risen sharply in the past 50 years. 

```{r medals_by_year, echo=FALSE, fig.width=12, message=FALSE, warning=FALSE}
#Get the initial data
medals_by_year <-
medals %>% 
  filter(medal != "") %>% 
  arrange(games)

#Arrange the data for the plot.
plot_medals_by_year <-
medals_by_year %>% 
  group_by(games, medal) %>% 
  summarise(medal_count = n()) %>% 
  mutate(season = if_else(str_detect(games, "Summer"),
                          "Summer",
                          "Winter"),
         year = as.numeric(str_extract_all(games, "[:digit:]{4}")))

#Draw the plot.
plot_medals_by_year %>% 
  mutate(medal = factor(medal, levels = c("Gold", "Silver","Bronze"))) %>% 
  ggplot(mapping=aes(x    =year, 
                     y    =medal_count,
                     color=medal))+
  geom_line()+
  facet_wrap(vars(season), scales = "free_y")+
  labs(title="Number of medals given out each year",
       x="", y="Number of medals")
```

