---
title: 'FDS Final Project: Report #2'
author: "Satu Iho"
output: html_document
always_allow_html: true
---

```{r setup, warning=FALSE, include=FALSE}
#Basic setup
library(httr) 
library(rvest)
library(stringr)
library(dplyr)
library(tidyr)
library(lubridate)
library(digest)
library(styler)
library(ggmap)
library(remedy)
library(janitor)
library(DBI) 
library(purrr)
library(xtable)
library(infer)
library(DT)
library(leaflet)
library(kableExtra)

zh_politicians_db <- dbConnect(RSQLite::SQLite(), "zh_politicians.db") 

dbListTables(zh_politicians_db)

#Run in the tables.
persons_tbl      <- tbl(zh_politicians_db, "PERSONS") 
addresses_tbl    <- tbl(zh_politicians_db, "ADDRESSES")
mandates_tbl     <- tbl(zh_politicians_db, "MANDATES") 
affiliations_tbl <- tbl(zh_politicians_db, "AFFILIATIONS")

#Convert the tables into tibbles.
persons      <- as_tibble(persons_tbl)
addresses    <- as_tibble(addresses_tbl)
mandates     <- as_tibble(mandates_tbl)
affiliations <- as_tibble(affiliations_tbl)

#Clean up the column names.
persons      <- clean_names(persons)
addresses    <- clean_names(addresses)
mandates     <- clean_names(mandates)
affiliations <- clean_names(affiliations)
```

### Part 1. Politicians with an active mandate  

The below plot displays the evolution of the number of Zürich politicians with 
an active mandate by assembly.   

```{r mandate_plot, echo=FALSE, fig.width=10, message=FALSE, warning=FALSE}

#First, prepare the data for the plot. Having checked the data there is one 
#entry with no start year and several with no end year, so I'll remove the one 
#with no start date and add the current year 2021 as the end date for the ones 
#with no end date yet.
mandates <-
mandates %>% 
  filter(mandate_start_year > 1800) %>%       
  mutate(mandate_end_year = replace(mandate_end_year, 
                                    mandate_end_year == 0, 
                                    2021))    

#Then, add years_active to indicate active mandate years 
mandate_evolution <-
  mandates %>% 
  mutate(years_active = map2(mandate_start_year, 
                             mandate_end_year, 
                             seq)) %>%             #Calculate the years active.
  select(person_id, assembly, years_active) %>%     #Exclude redundant columns.
  unnest()             #Unnest the years_active values to obtain a long tibble.


#Then, draw the plot.
mandate_evolution %>% 
  group_by(assembly, years_active) %>% 
  summarise("politicians_with_mandate"=n()) %>% 
  ggplot(mapping=aes(x    =years_active, 
                     y    =politicians_with_mandate,
                     color=assembly))+
  geom_line()+
  labs(title="Number of active mandates each year",
       subtitle="The peaks are caused by election years when multiple mandates 
       were active for the same seat.",
       x="Years", y="Number of mandates")
```

### Part 2. Faceted mandates  

This part shows the number of mandates in terms of assembly and gender. 

```{r mandate_faceted, echo=FALSE, fig.width=10, message=FALSE, warning=FALSE}
#Bring in the gender column from the persons table.  
mandates_with_gender <-
mandates %>% 
  left_join(persons, by=c("person_id"="id")) %>% 
  select(person_id, gender, assembly, mandate_start_year, mandate_end_year)

#Next, I can reuse the code from Part 1, with the addition of gender.
#First, clean up the data from outliers, null values etc.
mandates_with_gender_evolution <-
  mandates_with_gender %>% 
  filter(gender != "") %>%                #Remove entries with no gender value. 
  filter(mandate_start_year > 1800) %>%   #Remove the one entry without a start year.
  mutate(mandate_end_year = if_else(mandate_end_year == 0, 
                                    2021,
                                    mandate_end_year))    #Fill in 2021 as 
#mandate end year for the ones entries missing values (0).

#Then, add years_active and exclude unnecessary columns.
mandates_with_gender_evolution <-
  mandates_with_gender_evolution %>% 
  mutate(years_active = map2(mandate_start_year, 
                               mandate_end_year, 
                               seq)) %>%                
  unnest(years_active)  %>% 
  select(-starts_with("mandate"))                      

#Then, draw the plot itself with ggplot.
mandates_with_gender_evolution %>% 
  group_by(assembly, years_active, gender) %>% 
  summarise("politicians_with_mandate"=n()) %>% 
  ggplot(mapping=aes(    x = years_active, 
                         y = politicians_with_mandate,
                     color = gender))+
  geom_line()+
  facet_wrap(vars(assembly), scales = "free_y")+
  labs(title="Number of active mandates each year by assembly and gender.",
           x="Years", 
           y="Number of mandates")

```
The graphs paint a rather interesting picture in terms of women's participation
in Zürich politics: while non-existent until the start of the 1900's, the 
numbers seem to be converging for the Cantonal Council. 

### Part 3. Party affiliations in 2000.  

```{r parties_in_2000, echo=FALSE, message=FALSE, warning=FALSE}
#Use the affiliations tibble as the base.
affiliations_2000 <-
  affiliations %>% 
  filter(affiliation_start_year > 1800 & 
           affiliation_end_year > 1800 & 
           affiliation_end_year < 2022) %>%     #Include only complete records.
  select(person_id, party, ends_with("year")) %>%       
  mutate(affiliation_active = map2(affiliation_start_year, 
                                   affiliation_end_year, 
                                   seq)) %>%                 
  unnest() %>% 
  filter(affiliation_active == 2000)                        

#Add in the assembly details for the remaining rows.
affiliations_2000 <-
  affiliations_2000 %>% 
  inner_join(mandates, by="person_id") %>%      
  select(person_id, party, affiliation_active, assembly) %>% 
  filter(party != "")

#Then, draw the pie chart.
affiliations_2000 %>% 
  group_by(party, assembly) %>% 
  summarise("party_members"=n()) %>% 
  ggplot(aes(x="", y=party_members, fill=party)) +
  geom_bar(stat="identity", width=1, position = "fill") + 
  facet_wrap(vars(assembly))+
  coord_polar("y", start=0)+
  theme_void()

#And show the table
affiliations_2000 %>% 
  select(-affiliation_active) %>% 
  datatable()
```

As was also seen in the previous two graphs, only the Cantonal and Executive 
Councils were active in 2000, with the Cantonal Council being much larger (and 
diverse in terms of parties as shown above).

### Part 4. Showing historical party representation by assembly.
```{r party_historical, echo=FALSE, fig.width=10, message=FALSE, warning=FALSE}
#As in the previous section, use the affiliations table as the base. 
affiliations_evolution <-
affiliations %>% 
  filter(affiliation_start_year > 1800, 
         party != "", 
         affiliation_end_year > 1800,
         affiliation_end_year < 2022) %>%            
  select(person_id, party, ends_with("year")) %>%                   
  mutate(affiliation_active = map2(affiliation_start_year, 
                                   affiliation_end_year, 
                                   seq))   %>%                      
  unnest() %>% 
  select(-ends_with("year"))

#Add in the assembly details for the remaining rows.
affiliations_evolution <-
  affiliations_evolution %>% 
  inner_join(mandates, by="person_id") %>% 
  select(person_id, party, affiliation_active, assembly)

#The party name column contains a lot of what seem to be incorrect/nonsensical 
#entries, so the below code cleans them up.
affiliations_evolution <-
  affiliations_evolution %>% 
  mutate(party_corrected = case_when(                   
     #party == "AL" ~ "AL", 
     party == "BDP" ~ "BDP",
     #party == "BGB" ~ "Historical",
     party == "CSP" ~ "CSP",
     party == "CVP" ~ "CVP",
     party == "EDU" ~ "EDU",
     #party == "Evangelische Volkspartei" ~ "EVP",
     party == "EVP" ~ "EVP",
     party == "FDP" ~ "FDP",
     #party == "FPS" ~ "Minor",
     party == "GLP" ~ "GLP",
     party == "GP"  ~ "Grüne",
     party == "Grüne" ~ "Grüne",
     #party == "LdU" ~ "Historical",
     #party == "LS" ~ "AL",
     party == "PdA" ~ "PdA",
     #party == "POCH" ~ "Historical",
     party == "SD" ~ "SD",
     party == "SP" ~ "SP",
     party == "SVP" ~ "SVP",
     party == "SVP/BGB" ~ "SVP",
     str_detect(party, "Liste") ~ "List",  #Put all "Liste" ones into one group
     TRUE ~ "Unknown")) %>%                #Put all others in an unknown group
  filter(party_corrected != "Unknown")     #Remove the unknown entries

#Finally, draw the graph
affiliations_evolution %>% 
  filter(party_corrected != "List") %>% 
  group_by(party_corrected, assembly, affiliation_active) %>% 
  summarise("no_politicians"=n()) %>% 
  ggplot(mapping=aes(    x = affiliation_active, 
                         y = no_politicians,
                      fill = party_corrected))+
  geom_area()+
  facet_wrap(vars(assembly))+
  labs(title="Number of party members over time.",
       subtitle="Note: only the larger parties with clearly identifiable names
                 have been included.",
              x="Years", 
              y="Number of party members")
```
The above graphs show the number of party members per party and by assembly over
the years.   
*Note: only modern parties whose names were clearly identifiable as per [Wikipedia](https://en.wikipedia.org/wiki/List_of_political_parties_in_Switzerland) 
from the data set are included. The decreasing number of party members (which 
skews towards recent years) may be partially explained as politicians with no 
affiliation end date were excluded from the analysis. Excluding the incomplete 
records made the graph more easily readable which is why this strategy was 
chosen over for instance defaulting all records with no end year to e.g. 2021.* 

### Part 5. Politician life spans

This section analyses politicians life spans (in years).  

```{r life_span, include=FALSE}
#Prepare the data by including only complete records, converting to numeric and
#filtering out any potentially remaining nonsensical values.
persons_with_life_span <-
persons %>% 
  filter(year_of_birth > 1800 & year_of_death > 1800) %>% 
  mutate(year_of_birth = as.numeric(year_of_birth), 
         year_of_death = as.numeric(year_of_death)) %>% 
  mutate(life_span = year_of_death - year_of_birth) %>% 
  filter(life_span > 18 & life_span < 100)

#Save the average life_span into a variable to display within the output text. 
avg_life_span <-
persons_with_life_span %>% 
  summarise(count = n(),
            min_life_span = min(life_span),
            max_life_span = max(life_span),
            avg_life_span = mean(life_span)) %>% 
  pull(avg_life_span)

```
The average life span (in years) for politicians is 
**`r round(avg_life_span, digits=2)`**.  

The average life span (in years) for politicians according to title is displayed
below: 

```{r ages_titles, echo=FALSE, paged.print=FALSE, results='asis'}
#Use the same code as above but exclude records with blank titles.
persons_with_life_span_title <-
persons %>% 
  filter(year_of_birth > 1800 & year_of_death > 1800) %>% 
  mutate(year_of_birth = as.numeric(year_of_birth), 
         year_of_death = as.numeric(year_of_death)) %>% 
  mutate(life_span = year_of_death - year_of_birth) %>% 
  filter(life_span > 18 & life_span < 100 & title != "")

#Create reference tibble to view unique titles for harmonisation below.
unique_titles <-
persons_with_life_span_title %>% 
  select(title) %>% 
  unique()

#Harmonise the titles a little bit for more sensible analysis.
persons_with_life_span_title_corrected <-
persons_with_life_span_title %>% 
  mutate(title_corrected = case_when(                   
     str_detect(title, regex("prof", ignore_case = TRUE)) ~ "Prof.",
     str_detect(title, regex("dr",   ignore_case = TRUE)) ~ "Dr.",
     str_detect(title, regex("lic",  ignore_case = TRUE)) ~ "Lic.",
     TRUE ~ "Other")) %>% 
  select(id, life_span, title_corrected)

#Calculate average life span per corrected title.
avg_life_span_with_title <-
persons_with_life_span_title_corrected %>% 
  group_by(title_corrected) %>% 
  summarise(count = n(),
            avg_life_span = round(mean(life_span), 
                                  digits = 2))  %>% 
  rename("Title" = title_corrected,
         "Count" = count,
         "Average life span" = avg_life_span)

#Print the table with xtable.
avg_life_span_with_title %>% 
  kbl(caption = "Politician average life spans by title") %>% 
  kable_paper(full_width = F)
```
  
Next, we compare politicians with and without titles in terms of their lifespan.   

```{r life_span_title_vs_notitle, echo=FALSE, warning=FALSE}

#Apply the same title correction as above.
persons_with_life_span_title_categorised <-
persons_with_life_span  %>% 
  mutate(title_categorised = if_else(title == "", 
                                        "No title",
                                        "Title")) %>% 
  select(id, life_span, title_categorised)

#Then, draw a plot of the average age for the two groups.
persons_with_life_span_title_categorised %>% 
  group_by(title_categorised) %>% 
  summarise(avg_life_span = mean(life_span)) %>% 
  ggplot(mapping = aes(   x = title_categorised,
                          y = avg_life_span,
                       fill = title_categorised))+
  geom_col() +
  geom_text(aes(label = round(avg_life_span, digits=2)), 
            position = position_stack(vjust = .5))+
  scale_y_continuous(breaks = c(20, 30, 40, 50, 60, 70, 80))+
  theme(legend.position = "none")+ 
  labs(title="Average life span for politicians with no title and title",
           x="", 
           y="Average life span")
```

Based on the chart, it is obvious that there is a difference in the average life
span between the two groups, but whether it's big enough to be statistically 
significant or not is hard to say.

```{r t_test_lifespan, echo=FALSE, warning=FALSE}
#Perform t-test with the data created above.

t_test_p <-
persons_with_life_span_title_categorised %>% 
  t_test(formula = life_span ~ title_categorised,
                   order = c("No title", "Title"),
                   conf_level = 0.95) %>% 
  pull(p_value)
```
The t-test's p-value **`r t_test_p`** suggests that there is indeed a 
statistically significant difference between politician life spans depending on
if they have a title or not. 

### Part 6. Politicians life spans, also considering their birth before and 
### after 1918.

Similarly as the previous part, this part examines the life spans of politicians
with and without titles, but in addition distinguishes two subgroups: 
politicians born before and after 1918.  

```{r pre1918_vs_post1918, echo=FALSE, message=FALSE, warning=FALSE}
#Reuse the already cleansed tibble created at the start of Part 5 
persons_with_life_span_before_1918 <-
  persons_with_life_span%>% 
  mutate(born_before_1918 = year_of_birth < 1918,
         title_categorised = if_else(title == "", 
                                     "No title",
                                      "Title"))

#In preparation for the graph, change the facet labels to more sensible ones.
label_names_1918 <- c(`TRUE` = "Born before 1918",
                     `FALSE` = "Born after 1918")

#Finally, draw a faceted graph to show the two groups and their subgroups.
persons_with_life_span_before_1918 %>% 
  group_by(title_categorised, born_before_1918) %>% 
  summarise(avg_life_span = mean(life_span)) %>% 
  ggplot(mapping = aes(   x = title_categorised,
                          y = avg_life_span,
                       fill = title_categorised))+
  geom_col() +
  geom_text(aes(label = round(avg_life_span, digits=2)), 
            position = position_stack(vjust = .5))+
  scale_y_continuous(breaks = c(20, 30, 40, 50, 60, 70, 80))+
  theme(legend.position = "none")+ 
  facet_wrap(vars(born_before_1918), labeller = as_labeller(label_names_1918))+
  labs(title="Average life span for politicians with no title and title",
       subtitle = "Now also distinguishing between born after and before 1918",
           x="", 
           y="Average life span")

```

The graph suggests that the life span of politicians varies less depending on 
whether they have a title or not if they were born **after 1918** than if they 
were born **before 1918**. Accordingly, I would expect the t-test to show a 
statistically significant difference in the life span means for the group 
**born before 1918** only.

```{r t_test_before_1918, echo=FALSE}
#Do one t-test for the group born before 1918. 
t_persons_with_life_span_before_1918 <-
persons_with_life_span_before_1918 %>% 
  filter(born_before_1918 == TRUE) %>% 
  t_test(formula = life_span ~ title_categorised,
                   order = c("No title", "Title"),
                   conf_level = 0.95) 

t_persons_with_life_span_before_1918 %>% 
  kbl(caption = "t-test - politicians born before 1918") %>% 
  kable_paper(full_width = F)
```

This t-test result seems to confirm the initial expectation based on the graph,
namely that there is a statistically significant difference in the average life
span of a politician born before 1918, depending on if they had a title or not. 
```{r t_test_after_1918, echo=FALSE}
#Do one t-test for the group born after 1918. 
t_persons_with_life_span_after_1918 <-
persons_with_life_span_before_1918 %>% 
  filter(born_before_1918 == FALSE) %>% 
  t_test(formula = life_span ~ title_categorised,
                   order = c("No title", "Title"),
                   conf_level = 0.95) 

t_persons_with_life_span_after_1918 %>% 
  kbl(caption = "t-test - politicians born after 1918") %>% 
  kable_paper(full_width = F)
```

As expected, the hypothesis for this t-test cannot be rejected, suggesting that
there is no statistically significant difference in the average life span of 
politicians with and without titles if they were born after 1918. 

### Part 7. Top 10 politicians with mandates

The below graph shows the top ten politicians in terms of how many mandates they
have held.

```{r top_10_mandates, echo=FALSE, message=FALSE, warning=FALSE}

#First, find the 10 politicians with the most mandates.
mandates_top_10 <-
mandates %>% 
  count(person_id, sort=TRUE) %>% 
  head(10)

#Then, get the names added.
mandates_top_10 <-
mandates_top_10 %>% 
  inner_join(persons, by=c("person_id"="id")) %>% 
  select(person_id, n, ends_with("name")) %>% 
  mutate(name = str_glue("{firstname} {lastname}"))

#Then, draw the chart.
mandates_top_10 %>% 
 ggplot(mapping = aes(x = name,
                      y = n))+
  geom_col() +
  coord_flip()+
  labs(title="Top 10 number of mandates",
           x="", 
           y="Number of mandates")+
  theme_minimal()
 
```

### Part 8. Overlapping mandates

The below table displays politicians that have had overlapping mandates over the
years.  

```{r overlapping_mandates, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#First prepare the data, only including records where there is both genuine
#start and end dates for a mandate.
mandates_dates <-
  mandates %>% 
  filter(mandate_start_day > 1 & 
           mandate_start_month > 1 & 
           mandate_start_year > 1800,
         mandate_end_day > 1 & 
           mandate_end_month > 1 & 
           mandate_end_year > 1800) %>% 
  mutate(mandate_start_date = dmy(str_glue("{mandate_start_day}
                                           -{mandate_start_month}
                                           -{mandate_start_year}")),
         mandate_end_date   = dmy(str_glue("{mandate_end_day}
                                           -{mandate_end_month}
                                           -{mandate_end_year}"))) %>% 
  select(person_id, assembly, ends_with("date")) %>% 
  arrange(person_id, mandate_start_date)

#Create a filter to retain only records for persons with more than 1 mandate.
mandate_filter <-
  c(mandates_dates %>% 
      group_by(person_id) %>% 
      summarise(count = n()) %>% 
      filter(count >= 2) %>% 
      pull(person_id))

#Look for mandate overlaps with the lead function.
mandate_overlaps <-
  mandates_dates %>% 
  filter(person_id %in% mandate_filter) %>% 
  group_by(person_id) %>% 
  mutate(next_mandate_start_date = lead(mandate_start_date, n=1)) %>% 
  mutate(mandate_overlap = if_else(next_mandate_start_date > mandate_end_date, 
                                   FALSE, 
                                   TRUE)) %>% 
  filter(mandate_overlap == TRUE) 

#Save just the person id's with overlap into a tibble.
mandate_overlap_person_ids <-
  mandate_overlaps %>% 
  select(person_id) %>% 
  unique()

#Add in the names of the person id's identified above.
mandate_overlap_names <-
  mandate_overlap_person_ids %>% 
  left_join(persons, by=c("person_id"="id")) %>% 
  mutate(name = str_glue("{firstname} {lastname}")) %>% 
  select(name)

#Print the names only.
datatable(mandate_overlap_names)
```

### Part 9. Party affiliations changes over the years.  

The below table displays how party affiliations may have been switched for 
individual politicians over the years.

```{r overlapping_party_affiliations, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#First prepare the data, only including records where there is both genuine
#start and end dates for a party affiliation.
affiliation_dates <-
  affiliations %>% 
  filter(party != "",) %>% 
  filter(affiliation_start_day > 1 & 
           affiliation_start_month > 1 & 
           affiliation_start_year > 1800,
         affiliation_end_day > 1 & 
           affiliation_end_month > 1 & 
           affiliation_end_year > 1800 &
           affiliation_end_year < 2022) %>% 
  mutate(party_corrected = case_when(                   
    party == "BDP" ~ "BDP",
    party == "CSP" ~ "CSP",
    party == "CVP" ~ "CVP",
    party == "EDU" ~ "EDU",
    party == "EVP" ~ "EVP",
    party == "FDP" ~ "FDP",
    party == "GLP" ~ "GLP",
    party == "GP"  ~ "Grüne",
    party == "Grüne" ~ "Grüne",
    party == "PdA" ~ "PdA",
    party == "SD" ~ "SD",
    party == "SP" ~ "SP",
    party == "SVP" ~ "SVP",
    party == "SVP/BGB" ~ "SVP",
    str_detect(party, "Liste") ~ party,  #This time retain lists as they are
    TRUE ~ "Unknown")) %>%               
  filter(party_corrected != "Unknown") %>% 
  mutate(affiliation_start_date = dmy(str_glue("{affiliation_start_day}
                                               -{affiliation_start_month}
                                               -{affiliation_start_year}")),
         affiliation_end_date   = dmy(str_glue("{affiliation_end_day}
                                               -{affiliation_end_month}
                                               -{affiliation_end_year}"))) %>% 
  select(person_id, ends_with("date"), party_corrected) %>% 
  arrange(person_id)

#Look for politicians with more than one party affiliation.
affiliation_overlaps <-
  affiliation_dates %>% 
  group_by(person_id) %>% 
  mutate(party_different = lead(party_corrected, n=1)) %>% 
  mutate(party_overlap = if_else(party_corrected == party_different,
                                 FALSE, 
                                 TRUE)) %>% 
  filter(party_overlap == TRUE) 

#Store results for DT display
affiliation_overlaps_display <-
  affiliation_overlaps %>% 
  select(-starts_with("affiliation"), -party_overlap) %>% 
  rename("Person ID" = person_id, 
         "Party switched from" = party_corrected,
         "Party switched to" = party_different)

#Display results in DT table.
datatable(affiliation_overlaps_display)
```

Looking at the list of politicians that have switched parties over the years, it
seems that this was more common prior to the modern party era and rather 
politicians would switch between lists *(I know next to nothing about especially
historical Swiss politics so I'm not entirely sure what these Lists are 
though...)*  It can be noted that the only modern era politicians to have 
switched parties are politicians that have switched from Grüne to GLP.  

### Part 10. Politician address mapping (sample of 20)

The map below depicts the addresses of politicians selected randomly from the 
dataset.

```{r sample_addresses_map, echo=FALSE, message=FALSE, warning=FALSE}
#Create a tibble for names and locations to fetch long and lat data.
politician_addresses <-
  addresses %>% 
  left_join(persons, by=c("person_id"="id")) %>% 
  select(firstname, lastname, street, house_number, postal_code, city) %>% 
  filter(postal_code != "" &
           street != "") %>% 
  mutate(postal_code_corrected = str_extract_all(postal_code, "[:digit:]{4}",), 
         postal_code_corrected = as.character(postal_code_corrected),
         location = str_glue("{street} {house_number}, 
                              {postal_code_corrected}, 
                              {city}"),
         name = str_glue("{firstname} {lastname}")) %>% 
  select(name, location, everything())

#As in Report 1, get longitude and latitude data using the Google geolocation API
register_google(key = "AIzaSyD4KWAm6FqMs-5k8ARR-mgyIbYwzscYaRk")
getOption("ggmap")

#Store the lon-lat data of a sample of 20 addresses into a tibble.
politician_addresses_long_lat <-
  politician_addresses %>% 
  sample_n(20) %>%                    
  mutate_geocode(location)  

#Draw the map with Leaflet.
politician_addresses_long_lat %>%        
  leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng=~lon,lat=~lat,
             label=~name)      #Add politician name as a hover over label.
```


